
##Analyses and figures for Benkwitt, Wilson, & Graham "Seabird nutrient subsidies alter patterns of algal abundance and fish biomass on coral reefs following a bleaching event", Global Change Biology (In Review)

##Benthos


##load data
```{r, include = FALSE}
#load data
benthic<-read.csv("seabird_rat_benthic_data_2015_2018.csv", header=TRUE)
head(benthic)
```


##Sum percent of each category by transect
```{r, include = FALSE}
library(plyr)

#aggregate by transect
benthic_trans_group<-ddply(benthic, c("Year", "Island", "Transect", "Group"), summarise, 
                      Percent=sum(Percent), 
                      Proportion=sum(Proportion), 
                      .drop = FALSE)
benthic_trans_group
levels(benthic_trans_group$Island)

#drop islands not surveyed in 2018, since these values are actually NA's, not 0's
benthic_trans_group<-benthic_trans_group[!(benthic_trans_group$Year=="2018" & benthic_trans_group$Island=="Eagle") ,]
benthic_trans_group<-benthic_trans_group[!(benthic_trans_group$Year=="2018" & benthic_trans_group$Island=="Nelson_Island") ,]
nrow(benthic_trans_group)
benthic_trans_group
```


##Merge data with metadata
```{r, include = FALSE}
#load transect metadata to merge
uvc_meta <- read.csv("seabird_rat_transect_metadata.csv", header=TRUE)
head(uvc_meta)

##merge dataframes so have transect metadata with benthic data
benthic_trans_group_meta<-merge(uvc_meta, benthic_trans_group, c("Year", "Island", "Transect"))
benthic_trans_group_meta
```


##NMDS
```{r}
library(reshape2)
library(vegan)

#Aggregate to island-level
benthic_island_group <-ddply(benthic_trans_group_meta, c("Year","Treatment", "Atoll", "Island", "Group"), summarise,
          Mean_Proportion = mean(Proportion), 
          sd = sd(Proportion),
          n = length(Proportion),
          se = sd(Proportion)/sqrt(n))
benthic_island_group

#Change year to factor
benthic_island_group$Year<-as.factor(benthic_island_group$Year)

#Add column for combined treatment and year (to make some comparisons and plotting easier)
benthic_island_group$Treat_Year<-paste(benthic_island_group$Treatment, benthic_island_group$Year, sep = '_')
#change to factor
benthic_island_group$Treat_Year<-as.factor(benthic_island_group$Treat_Year)


#convert to wide for NMDS
benthic_island_wide <- dcast(benthic_island_group, Year + Atoll + Island + Treatment + Treat_Year ~ Group, value.var="Mean_Proportion")
benthic_island_wide

#run NMDS
nmds_benth<-metaMDS(benthic_island_wide[6:ncol(benthic_island_wide)], distance = "bray", trymax=200, k=2)
nmds_benth
scores(nmds_benth, display="species")
#stress = 0.11 (good)

#look at plot and centroids
plot(nmds_benth)
ord.fit<-envfit(nmds_benth~Treat_Year, data = benthic_island_wide)
ord.fit
plot(ord.fit)
#draw polygons
ordihull(nmds_benth, benthic_island_wide$Treat_Year, pty=3)
#looks like some differences among groups, now run statistics

#PERMANOVA
#anosim more influenced by dispersion, so use adonis

#set blocks by atoll
perm <- how(nperm = 999)
setBlocks(perm) <- with(benthic_island_wide, Atoll)

#run analysis
benthic_ad<-adonis2(benthic_island_wide[6:ncol(benthic_island_wide)] ~ Treatment*Year, data = benthic_island_wide, permutations = perm)
benthic_ad
#interaction significant
#treatment and year both significant
#(note: permutation test, so slightly different p-values each time)


#test dispersion using PERMDISP
benthic_disp <- betadisper(vegdist(benthic_island_wide[6:ncol(benthic_island_wide)], method = "bray"), benthic_island_wide$Treat_Year)
benthic_disp

#test for overall differences
anova(benthic_disp)

## Permutation test for pairwise comparisons
permutest(benthic_disp, pairwise = TRUE)
#(note: permutation test, so slightly different p-values each time)


#extract mean dispersion values (distance to centroid), along with se's and 95% CI
mod_benth.mean<-tapply(benthic_disp$distances, benthic_island_wide$Treat_Year, mean)
mod_benth.mean
mod_benth.sd<-tapply(benthic_disp$distances, benthic_island_wide$Treat_Year, sd)
mod_benth.length<-tapply(benthic_disp$distances, benthic_island_wide$Treat_Year, length)
mod_benth.se<-mod_benth.sd/sqrt(mod_benth.length)
mod_benth.se
mod_benth.ci_low<-mod_benth.mean-(1.96*mod_benth.se)
mod_benth.ci_high<-mod_benth.mean+(1.96*mod_benth.se)
#combine into a dataframe
mod_benth.out<-as.data.frame(cbind(mod_benth.mean,mod_benth.se, mod_benth.ci_low, mod_benth.ci_high ))
mod_benth.out<-cbind(Treat_Year = rownames(mod_benth.out), mod_benth.out)
mod_benth.out


##SIMPER analysis to test effects of species
sim_benth<-simper(benthic_island_wide[6:ncol(benthic_island_wide)], benthic_island_wide$Treat_Year, permutations = 999)
summary(sim_benth)

```

##Extract NMDS results and plot
```{r}
library(ggplot2)

##Extract site and nmds_spp
spp.sc <- scores(nmds_benth, display = "species", shrink = FALSE) 
spp.sc
site.sc <- scores(nmds_benth, display = "sites", shrink = FALSE) 
site.sc

#merge site.sc with metadata from dataframe
benth_island_nmds<-cbind(site.sc, benthic_island_wide[1:5])
benth_island_nmds

#extract species scores
species.scores <- as.data.frame(scores(nmds_benth, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
head(species.scores)  #look at the data


#get convex hulls for polygons
hull_nr2015<-benth_island_nmds[benth_island_nmds$Treat_Year == "No_rats_2015", ][chull(benth_island_nmds[benth_island_nmds$Treat_Year == "No_rats_2015", ]$NMDS1, benth_island_nmds[benth_island_nmds$Treat_Year == "No_rats_2015", ]$NMDS2), ] 
hull_nr2018<-benth_island_nmds[benth_island_nmds$Treat_Year == "No_rats_2018", ][chull(benth_island_nmds[benth_island_nmds$Treat_Year == "No_rats_2018", ]$NMDS1, benth_island_nmds[benth_island_nmds$Treat_Year == "No_rats_2018", ]$NMDS2), ] 
hull_r2018<-benth_island_nmds[benth_island_nmds$Treat_Year == "Rats_2018", ][chull(benth_island_nmds[benth_island_nmds$Treat_Year == "Rats_2018", ]$NMDS1, benth_island_nmds[benth_island_nmds$Treat_Year == "Rats_2018", ]$NMDS2), ] 
hull_r2015<-benth_island_nmds[benth_island_nmds$Treat_Year == "Rats_2015", ][chull(benth_island_nmds[benth_island_nmds$Treat_Year == "Rats_2015", ]$NMDS1, benth_island_nmds[benth_island_nmds$Treat_Year == "Rats_2015", ]$NMDS2), ] 
hull.data <- rbind(hull_nr2015, hull_nr2018, hull_r2018, hull_r2015)  #combine grp.a and grp.b
hull.data

#get centroids to plot arrows
ord.fit<-envfit(nmds_benth~Treat_Year, data = benthic_island_wide)
ord.fit


nmds_b<-
ggplot() + 
 geom_polygon(data=hull.data[hull.data$Treatment=="Rats",],aes(x=NMDS1,y=NMDS2,fill=Treatment,group=Treat_Year),colour = 'black', alpha = 0.15) + # add the convex hulls - two steps because the blue is fainter - alpha=0.15, 
   geom_polygon(data=hull.data[hull.data$Treatment=="No_rats",],aes(x=NMDS1,y=NMDS2,fill=Treatment,group=Treat_Year), colour = 'black', alpha = 0.2) + # add the convex hulls - two steps because the blue is fainter - alpha=0.2,
  geom_text(data=species.scores,aes(x=NMDS1,y=NMDS2,label=species)) +  # add the species labels - 
  geom_point(data=benth_island_nmds,aes(x=NMDS1,y=NMDS2,shape=Year,colour=Treatment, fill = Treatment), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                           labels=c("Seabirds", "Rats") ) +
    geom_segment(aes(x = 0.2073, y = -0.0475, xend = 0.1407, yend = -0.1338), colour = "#e41a1c", size = 1.2, arrow =   arrow(length = unit(0.1, "inches")))+ #rats - from ord.fit
  geom_segment(aes(x = 0.1593, y = 0.0552, xend = -0.5806, yend = 0.1245), colour = "#56B4E9", size = 1.2, arrow =   arrow(length = unit(0.2, "inches")))+ #norats - from ord.fit
  #coord_equal() +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none")
        #legend.title = element_blank()) #remove legend title
        #legend.box.background = element_rect(colour = "black", fill='white', linetype='solid', size=1), 
       # legend.justification=c(1,.9), legend.position=c(1,.9))  #  fix legend
      
nmds_b


##Make a plot of dispersion means +/- SE
mod.out #from betadisper, above
#create new columns for Treament and year
mod.out$Treatment<-c("No_rats","No_rats","Rats", "Rats")
mod.out$Year<-c("2015","2018","2015", "2018")
mod.out

bend<-ggplot(mod.out, aes(x=Year, y=mod_benth.mean, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(mod_benth.mean-mod_benth.se), ymax=(mod_benth.mean+mod_benth.se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("(a) Benthos") +
  theme_bw() + 
    theme(#axis.text.y = element_blank(), #remove y-axis labels
  panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Dispersion", limits = c(0,0.4), breaks=seq(0,0.4,0.1)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching"))  

bend

```


#LMMs for individual groups
```{r}
library(lme4) #glmms

#look at data
benthic_trans_group_meta

#change year to factor
benthic_trans_group_meta$Year<-as.factor(benthic_trans_group_meta$Year)

#run final models, log-transformed when necessary to meet assumptions
#hard coral
coral_lmer_1_int<-lmer(Percent~Treatment*Year + (1|Atoll/Island), data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Hard_Coral",])
summary(coral_lmer_1_int)

coral_lmer_1_add<-lmer(Percent~Treatment+Year + (1|Atoll/Island),  data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Hard_Coral",])
summary(coral_lmer_1_add)

#LRT
anova(coral_lmer_1_int, coral_lmer_1_add)
#no sig interaction

#test for significance of each fixed effect with LRT's
coral_lmer_1_yr<-lmer(Percent~Treatment + (1|Atoll/Island),  data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Hard_Coral",])
coral_lmer_1_treat<-lmer(Percent~Year + (1|Atoll/Island),  data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Hard_Coral",])

anova(coral_lmer_1_add, coral_lmer_1_yr)
#very significant effect of year

anova(coral_lmer_1_add, coral_lmer_1_treat)
#no significant effect of treatment

#extract effect size and 95% CI
summary(coral_lmer_1_add)
confint(coral_lmer_1_add, method = "Wald")


#Halimeda
algae_lmer_int<-lmer(log(Percent + 1)~Treatment*Year + (1|Atoll/Island), data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Halimeda",])
summary(algae_lmer_int)

algae_lmer_add<-lmer(log(Percent + 1)~Treatment+Year + (1|Atoll/Island),data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Halimeda",])
summary(algae_lmer_add)

anova(algae_lmer_int, algae_lmer_add)
#very significant interaction

#extract effects and back-transform to original scale
algae_lmer_int<-lmer(log(Percent + 1)~relevel(Treatment, "Rats")*relevel(Year, "2018") + (1|Atoll/Island), data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Halimeda",])
summary(algae_lmer_int)
confint(algae_lmer_int, method = "Wald")
exp(fixef(algae_lmer_int)) #back to original scale (multiplicative change)
exp(confint(algae_lmer_int, method = "Wald"))#back to original scale (multiplicative change)


#CCA
cca_lmer_int<-lmer(log(Percent + 1)~Treatment*Year + (1|Atoll/Island), data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="CCA",])
summary(cca_lmer_int)

cca_lmer_add<-lmer(log(Percent + 1)~Treatment+Year + (1|Atoll/Island),  data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="CCA",])
summary(cca_lmer_add)

anova(cca_lmer_int, cca_lmer_add)
#interaction significant

#extract effects and back-transform
cca_lmer_int<-lmer(log(Percent + 1)~relevel(Treatment, "No_rats")*relevel(Year, "2018") + (1|Atoll/Island), data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="CCA",])
summary(cca_lmer_int)
confint(cca_lmer_int, method = "Wald")
exp(fixef(cca_lmer_int)) #back to original scale (multiplicative change)
exp(confint(cca_lmer_int, method = "Wald"))#back to original scale (multiplicative change)


#Pavement
pave_lmer_int<-lmer(Percent~Treatment*Year + (1|Atoll/Island), data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Pavement",])
summary(pave_lmer_int)

pave_lmer_add<-lmer(Percent~Treatment+Year + (1|Atoll/Island), data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Pavement",])
summary(pave_lmer_add)

anova(pave_lmer_int, pave_lmer_add)
#very significant year*treatment interaction

#extract effects
pave_lmer_int<-lmer(Percent~relevel(Treatment, "Rats")*relevel(Year, "2018") + (1|Atoll/Island), data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Pavement",])

summary(pave_lmer_int)
confint(pave_lmer_int, method = "Wald")#year


#Rubble
rub_lmer_int<-lmer(log(Percent + 1)~Treatment*Year + (1|Atoll/Island), data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Rubble",])
summary(rub_lmer_int)

rub_lmer_add<-lmer(log(Percent + 1)~Treatment+Year + (1|Atoll/Island), data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Rubble",])
summary(rub_lmer_add)

anova(rub_lmer_int, rub_lmer_add)
#no significant interaction

rub_lmer_yr<-lmer(log(Percent + 1)~Treatment + (1|Atoll/Island), data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Rubble",])

rub_lmer_treat<-lmer(log(Percent + 1)~Year + (1|Atoll/Island), data = benthic_trans_group_meta[benthic_trans_group_meta$Group=="Rubble",])


anova(rub_lmer_add, rub_lmer_yr)
anova(rub_lmer_add, rub_lmer_treat)
#neither year nor treatment significant

#extract effects
summary(rub_lmer_add)
confint(rub_lmer_add, method = "Wald")
exp(fixef(rub_lmer_add)) #back to original scale (multiplicative change)
exp(confint(rub_lmer_add, method = "Wald"))#back to original scale (multiplicative change)

```


##LMM for structural complexity
```{r}

#check meta-data
uvc_meta
#make year a factor
uvc_meta$Year<-as.factor(uvc_meta$Year)

#look at summarized data
structure_treat <-ddply(uvc_meta, c("Year","Treatment"), summarise,
          Mean_Structure = mean(Structure), 
          sd = sd(Structure),
          n = length(Structure),
          se = sd(Structure)/sqrt(n))

structure_treat
#looks similar across groups

#Run a model
sc_lmer_int<-lmer(Structure~Treatment*Year + (1|Atoll/Island), data = uvc_meta)
summary(sc_lmer_int)

sc_lmer_add<-lmer(Structure~Treatment+Year + (1|Atoll/Island), data = uvc_meta)
summary(sc_lmer_add)

anova(sc_lmer_int, sc_lmer_add)
#no significant interaction

sc_lmer_treat<-lmer(Structure~Year + (1|Atoll/Island), data = uvc_meta)
sc_lmer_yr<-lmer(Structure~Treatment + (1|Atoll/Island), data = uvc_meta)

anova(sc_lmer_add, sc_lmer_treat)
anova(sc_lmer_add, sc_lmer_yr)
#neither year nor treatment significant


#extract effects
sc_lmer_int<-lmer(Structure~relevel(Treatment, "No_rats")*relevel(Year, "2018") + (1|Atoll/Island), data = uvc_meta)
summary(sc_lmer_int)
confint(sc_lmer_int, method = "Wald")
```


##Plots of individual benthic groups
```{r, echo = FALSE}
library(ggplot2)
library('gridExtra') 

#get means and se's
benthic_group_treat <-ddply(benthic_trans_group_meta, c("Year","Treatment", "Group"), summarise,
          Mean_Percent = mean(Percent), 
          se = sd(Percent)/sqrt(length(Percent)))
benthic_group_treat
benthic_group_treat$Year<-as.factor(benthic_group_treat$Year)


###plot each FG seperately, then mix together
a<-ggplot(benthic_group_treat[benthic_group_treat$Group=="Hard_Coral",], aes(x=Year, y=Mean_Percent, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Percent-se), ymax=(Mean_Percent+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
   scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("(a) Hard coral") +
  theme_bw() + 
   theme( #axis.text.y = element_blank(), #remove y-axis labels
panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position = "none") + 
        scale_y_continuous(name = "Percent cover", limits = c(0,35), breaks=seq(0,30,10)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching"))  

b<-ggplot(benthic_group_treat[benthic_group_treat$Group=="CCA",], aes(x=Year, y=Mean_Percent, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Percent-se), ymax=(Mean_Percent+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("(b) CCA") +
  theme_bw() + 
    theme( #axis.text.y = element_blank(), #remove y-axis labels
panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        #legend.title = element_blank(), #remove legend title
        #legend.box.background = element_rect(colour = "black", fill='white', linetype='solid', size=1), 
       # legend.justification=c(1,.9), legend.position=c(1,.9)) + #  fix legend
        scale_y_continuous(name = "Percent cover", limits = c(0,35), breaks=seq(0,30,10)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching"))  


c<-ggplot(benthic_group_treat[benthic_group_treat$Group=="Halimeda",], aes(x=Year, y=Mean_Percent, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Percent-se), ymax=(Mean_Percent+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("(c) Halimeda") +
  theme_bw() + 
    theme(#axis.text.y = element_blank(), #remove y-axis labels
  panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Percent cover", limits = c(0,35), breaks=seq(0,30,10)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching"))  
        


d<-ggplot(benthic_group_treat[benthic_group_treat$Group=="Pavement",], aes(x=Year, y=Mean_Percent, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Percent-se), ymax=(Mean_Percent+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                           labels=c("Seabirds", "Rats") ) +
  ggtitle("(d) Pavement") +
  theme_bw() + 
  theme(#axis.text.y = element_blank(), #remove y-axis labels
  panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Percent cover", limits = c(0,56), breaks=seq(0,55,10)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching")) 

e<-ggplot(benthic_group_treat[benthic_group_treat$Group=="Rubble",], aes(x=Year, y=Mean_Percent, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Percent-se), ymax=(Mean_Percent+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                           labels=c("Seabirds", "Rats") ) +
  ggtitle("(e) Rubble") +
  theme_bw() + 
  theme(#axis.text.y = element_blank(), #remove y-axis labels
          panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Percent cover", limits = c(0,15), breaks=seq(0,15,5)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching")) 

#now plot structural complexity
structure_treat

f<-ggplot(structure_treat, aes(x=Year, y=Mean_Structure, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Structure-se), ymax=(Mean_Structure+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                           labels=c("Seabirds", "Rats") ) +
  ggtitle("(f) Structural complexity") +
  theme_bw() + 
  theme(#axis.text.y = element_blank(), #remove y-axis labels
          panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Rugosity index", limits = c(1,6), breaks=seq(1,6,1)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching")) 
f

#show plots
grid.arrange(a, b,c,d,e,f, nrow = 3) 

#export
g <- arrangeGrob(a, b,c,d,e,f, nrow = 3) #generates g
#ggsave("filename.eps", g, units="mm", width=180, height=270, dpi=1200)

```

#Look at hard coral by genus
```{r}

#make new data_frame with only hard coral
head(benthic)
coral_only<-benthic[benthic$Group=="Hard_Coral",]
head(coral_only)
#drop levels
coral_only$Species<-factor(coral_only$Species)
levels(coral_only$Species)

#sum each genus by transect
coral_sp_trans <-ddply(coral_only, c("Year","Island", "Transect", "Species"), summarise,
          Sum_Percent = sum(Percent), 
          .drop=FALSE)
coral_sp_trans

#merge with meta-data
coral_sp_trans_meta<-merge(uvc_meta, coral_sp_trans, c("Year", "Island", "Transect"))

#look at means by treatment and year
coral_sp_treat <-ddply(coral_sp_trans_meta, c("Year","Treatment", "Species"), summarise,
          Mean_Percent = mean(Sum_Percent), 
          sd = sd(Sum_Percent),
          n = length(Sum_Percent),
          se = sd(Sum_Percent)/sqrt(n),
          .drop=FALSE)
coral_sp_treat

#look at means by year
coral_sp_year <-ddply(coral_sp_trans_meta, c("Year","Species"), summarise,
          Mean_Prop = mean(Sum_Percent), 
          sd = sd(Sum_Percent),
          n = length(Sum_Percent),
          se = sd(Sum_Percent)/sqrt(n),
          .drop=FALSE)
coral_sp_year
#Acropora is dominant, and similar between birdy and ratty islands within years

#run stats on Acropra declines
acrop_int<-lmer(Sum_Percent~Treatment*Year + (1|Atoll/Island), data = coral_sp_trans_meta[coral_sp_trans_meta$Species=="Acropora",])
summary(acrop_int)

acrop_add<-lmer(Sum_Percent~Treatment+Year + (1|Atoll/Island), data = coral_sp_trans_meta[coral_sp_trans_meta$Species=="Acropora",])

anova(acrop_int, acrop_add)
#no sig interaction

acrop_treat<-lmer(Sum_Percent~Year + (1|Atoll/Island), data = coral_sp_trans_meta[coral_sp_trans_meta$Species=="Acropora",])

acrop_year<-lmer(Sum_Percent~Treatment + (1|Atoll/Island), data = coral_sp_trans_meta[coral_sp_trans_meta$Species=="Acropora",])

anova(acrop_add, acrop_treat)
#no sig effect of treatment

anova(acrop_add, acrop_year)
#very sig effect of year

#extract effects
summary(acrop_add)
confint(acrop_add, method = "Wald")
```