
##Analyses and figures for Benkwitt, Wilson, & Graham (2019) "Seabird nutrient subsidies alter patterns of algal abundance and fish biomass on coral reefs following a bleaching event" Global Change Biology 25:2619-2632.

##Benthos

##load data
```{r, include = FALSE}
#load data
benthic<-read.csv("seabird_rat_benthic_data_2015_2018.csv", header=TRUE)
head(benthic)
```


##Sum percent of each category by transect
```{r, include = FALSE}
library(plyr)

#aggregate by transect
benthic_trans_group<-ddply(benthic, c("Year", "Island", "Transect", "Group"), summarise, 
                      Percent=sum(Percent), 
                      Proportion=sum(Proportion), 
                      .drop = FALSE)
benthic_trans_group
levels(benthic_trans_group$Island)

#drop islands not surveyed in 2018, since these values are actually NA's, not 0's
benthic_trans_group<-benthic_trans_group[!(benthic_trans_group$Year=="2018" & benthic_trans_group$Island=="Eagle") ,]
benthic_trans_group<-benthic_trans_group[!(benthic_trans_group$Year=="2018" & benthic_trans_group$Island=="Nelson_Island") ,]
nrow(benthic_trans_group)
benthic_trans_group
```


##Merge data with metadata
```{r, include = FALSE}
#load transect metadata to merge
uvc_meta <- read.csv("seabird_rat_transect_metadata.csv", header=TRUE)
head(uvc_meta)

##merge dataframes so have transect metadata with benthic data
benthic_trans_group_meta<-merge(uvc_meta, benthic_trans_group, c("Year", "Island", "Transect"))
benthic_trans_group_meta
```


##NMDS
```{r}
library(reshape2)
library(vegan)

#Aggregate to island-level
benthic_island_group <-ddply(benthic_trans_group_meta, c("Year","Treatment", "Atoll", "Island", "Group"), summarise,
          Mean_Proportion = mean(Proportion), 
          sd = sd(Proportion),
          n = length(Proportion),
          se = sd(Proportion)/sqrt(n))
benthic_island_group

#Change year to factor
benthic_island_group$Year<-as.factor(benthic_island_group$Year)

#Add column for combined treatment and year (to make some comparisons and plotting easier)
benthic_island_group$Treat_Year<-paste(benthic_island_group$Treatment, benthic_island_group$Year, sep = '_')
#change to factor
benthic_island_group$Treat_Year<-as.factor(benthic_island_group$Treat_Year)


#convert to wide for NMDS
benthic_island_wide <- dcast(benthic_island_group, Year + Atoll + Island + Treatment + Treat_Year ~ Group, value.var="Mean_Proportion")
benthic_island_wide

#run NMDS
nmds_benth<-metaMDS(benthic_island_wide[6:ncol(benthic_island_wide)], distance = "bray", trymax=200, k=2)
nmds_benth
scores(nmds_benth, display="species")
#stress = 0.11 (good)

#look at plot and centroids
plot(nmds_benth)
ord.fit<-envfit(nmds_benth~Treat_Year, data = benthic_island_wide)
ord.fit
plot(ord.fit)
#draw polygons
ordihull(nmds_benth, benthic_island_wide$Treat_Year, pty=3)
#looks like some differences among groups, now run statistics

#PERMANOVA
#anosim more influenced by dispersion, so use adonis

#set blocks by atoll
perm <- how(nperm = 999)
setBlocks(perm) <- with(benthic_island_wide, Atoll)

#run analysis
benthic_ad<-adonis2(benthic_island_wide[6:ncol(benthic_island_wide)] ~ Treatment*Year, data = benthic_island_wide, permutations = perm)
benthic_ad
#interaction significant
#treatment and year both significant
#(note: permutation test, so slightly different p-values each time)


#test dispersion using PERMDISP
benthic_disp <- betadisper(vegdist(benthic_island_wide[6:ncol(benthic_island_wide)], method = "bray"), benthic_island_wide$Treat_Year)
benthic_disp

#test for overall differences
anova(benthic_disp)

## Permutation test for pairwise comparisons
permutest(benthic_disp, pairwise = TRUE)
#(note: permutation test, so slightly different p-values each time)


#extract mean dispersion values (distance to centroid), along with se's and 95% CI
mod_benth.mean<-tapply(benthic_disp$distances, benthic_island_wide$Treat_Year, mean)
mod_benth.mean
mod_benth.sd<-tapply(benthic_disp$distances, benthic_island_wide$Treat_Year, sd)
mod_benth.length<-tapply(benthic_disp$distances, benthic_island_wide$Treat_Year, length)
mod_benth.se<-mod_benth.sd/sqrt(mod_benth.length)
mod_benth.se
mod_benth.ci_low<-mod_benth.mean-(1.96*mod_benth.se)
mod_benth.ci_high<-mod_benth.mean+(1.96*mod_benth.se)
#combine into a dataframe
mod_benth.out<-as.data.frame(cbind(mod_benth.mean,mod_benth.se, mod_benth.ci_low, mod_benth.ci_high ))
mod_benth.out<-cbind(Treat_Year = rownames(mod_benth.out), mod_benth.out)
mod_benth.out


##SIMPER analysis to test effects of species
sim_benth<-simper(benthic_island_wide[6:ncol(benthic_island_wide)], benthic_island_wide$Treat_Year, permutations = 999)
summary(sim_benth)


##SIMPER analysis to test effects of species on atoll
sim_benth2<-simper(benthic_island_wide[6:ncol(benthic_island_wide)], benthic_island_wide$Atoll, permutations = 999)
summary(sim_benth2)

```

##Extract NMDS results and plot
```{r}
library(ggplot2)

##Extract site and nmds_spp
spp.sc <- scores(nmds_benth, display = "species", shrink = FALSE) 
spp.sc
site.sc <- scores(nmds_benth, display = "sites", shrink = FALSE) 
site.sc

#merge site.sc with metadata from dataframe
benth_island_nmds<-cbind(site.sc, benthic_island_wide[1:5])
benth_island_nmds

#extract species scores
species.scores <- as.data.frame(scores(nmds_benth, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
head(species.scores)  #look at the data


#get convex hulls for polygons
hull_nr2015<-benth_island_nmds[benth_island_nmds$Treat_Year == "No_rats_2015", ][chull(benth_island_nmds[benth_island_nmds$Treat_Year == "No_rats_2015", ]$NMDS1, benth_island_nmds[benth_island_nmds$Treat_Year == "No_rats_2015", ]$NMDS2), ] 
hull_nr2018<-benth_island_nmds[benth_island_nmds$Treat_Year == "No_rats_2018", ][chull(benth_island_nmds[benth_island_nmds$Treat_Year == "No_rats_2018", ]$NMDS1, benth_island_nmds[benth_island_nmds$Treat_Year == "No_rats_2018", ]$NMDS2), ] 
hull_r2018<-benth_island_nmds[benth_island_nmds$Treat_Year == "Rats_2018", ][chull(benth_island_nmds[benth_island_nmds$Treat_Year == "Rats_2018", ]$NMDS1, benth_island_nmds[benth_island_nmds$Treat_Year == "Rats_2018", ]$NMDS2), ] 
hull_r2015<-benth_island_nmds[benth_island_nmds$Treat_Year == "Rats_2015", ][chull(benth_island_nmds[benth_island_nmds$Treat_Year == "Rats_2015", ]$NMDS1, benth_island_nmds[benth_island_nmds$Treat_Year == "Rats_2015", ]$NMDS2), ] 
hull.data <- rbind(hull_nr2015, hull_nr2018, hull_r2018, hull_r2015)  #combine grp.a and grp.b
hull.data

#get centroids to plot arrows
ord.fit<-envfit(nmds_benth~Treat_Year, data = benthic_island_wide)
ord.fit


nmds_b<-
ggplot() + 
 geom_polygon(data=hull.data[hull.data$Treatment=="Rats",],aes(x=NMDS1,y=NMDS2,fill=Treatment,group=Treat_Year),colour = 'black', alpha = 0.15) + # add the convex hulls - two steps because the blue is fainter - alpha=0.15, 
   geom_polygon(data=hull.data[hull.data$Treatment=="No_rats",],aes(x=NMDS1,y=NMDS2,fill=Treatment,group=Treat_Year), colour = 'black', alpha = 0.2) + # add the convex hulls - two steps because the blue is fainter - alpha=0.2,
  geom_text(data=species.scores,aes(x=NMDS1,y=NMDS2,label=species)) +  # add the species labels - 
  geom_point(data=benth_island_nmds,aes(x=NMDS1,y=NMDS2,shape=Year,colour=Treatment, fill = Treatment), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                           labels=c("Seabirds", "Rats") ) +
    geom_segment(aes(x = 0.2073, y = -0.0475, xend = 0.1407, yend = -0.1338), colour = "#e41a1c", size = 1.2, arrow =   arrow(length = unit(0.1, "inches")))+ #rats - from ord.fit
  geom_segment(aes(x = 0.1593, y = 0.0552, xend = -0.5806, yend = 0.1245), colour = "#56B4E9", size = 1.2, arrow =   arrow(length = unit(0.2, "inches")))+ #norats - from ord.fit
  #coord_equal() +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none")
        #legend.title = element_blank()) #remove legend title
        #legend.box.background = element_rect(colour = "black", fill='white', linetype='solid', size=1), 
       # legend.justification=c(1,.9), legend.position=c(1,.9))  #  fix legend
      
nmds_b


##Make a plot of dispersion means +/- SE
mod.out #from betadisper, above
#create new columns for Treament and year
mod.out$Treatment<-c("No_rats","No_rats","Rats", "Rats")
mod.out$Year<-c("2015","2018","2015", "2018")
mod.out

bend<-ggplot(mod.out, aes(x=Year, y=mod_benth.mean, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(mod_benth.mean-mod_benth.se), ymax=(mod_benth.mean+mod_benth.se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("(a) Benthos") +
  theme_bw() + 
    theme(#axis.text.y = element_blank(), #remove y-axis labels
  panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Dispersion", limits = c(0,0.4), breaks=seq(0,0.4,0.1)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching"))  

bend

```


#LMMs for individual groups
```{r}
library(lme4) #lmms
library(blme)
library(ggplot2)
library(jtools) #plot_summ
library(MuMIn) #r.squaredGLMM

#look at data
benthic_trans_group_meta

#change year to factor
benthic_trans_group_meta$Year<-as.factor(benthic_trans_group_meta$Year)
#relevel factor
benthic_trans_group_meta$Treatment<-relevel(benthic_trans_group_meta$Treatment, "Rats")
#create column for logged biomass to make coding easier
benthic_trans_group_meta$lpercent<-log(benthic_trans_group_meta$Percent+1)

#run final models, log-transformed when necessary to meet assumptions

#hard coral
coral_blmer_int<-blmer(Percent~Treatment*Year  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Hard_Coral",])
summary(coral_blmer_int)

coral_blmer_add<-blmer(Percent~Treatment + Year+ (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Hard_Coral",])
summary(coral_blmer_add)

anova(coral_blmer_int, coral_blmer_add)
#no sig interaction

#test for sig of fixed effects
coral_blmer_add_treat<-blmer(Percent~Year  +  (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Hard_Coral",])

coral_blmer_add_year<-blmer(Percent~Treatment  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Hard_Coral",])

#year
anova(coral_blmer_add, coral_blmer_add_year)
#very sig decline by year

#treat
anova(coral_blmer_add, coral_blmer_add_treat)
#ns treatment effect

##test for sig of random effects
coral_blmer_null<-blmer(Percent~1  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Hard_Coral",])
anova(coral_blmer_int, coral_blmer_null)
##fixed effects very important

#extract effect size and 95% CI
coralsumm_add<-summ(coral_blmer_add, conf.method="profile",  confint=TRUE,scale=TRUE, pvals=FALSE)
coralsumm_add

coralsumm_int<-summ(coral_blmer_int, conf.method="profile",  confint=TRUE,scale=TRUE,  pvals=FALSE)
coralsumm_int

a<-plot_coefs(coralsumm_int, inner_ci_level=0.75, ci_levels=0.95, color.class=c("black", "black", "black"), conf.method="profile", coefs=c("Treatment:Year","Treatment","Year"))+ 
  ggtitle("Hard coral")+
  xlim(c(-16, 16))+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank()) #remove gridlines
a


r.squaredGLMM(object=coral_blmer_int)
#           R2m       R2c
# 0.09269879 0.6959157



#halimeda
halimeda_blmer_int<-blmer(lpercent~Treatment*Year  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Halimeda",])
summary(halimeda_blmer_int)

halimeda_blmer_add<-blmer(lpercent~Treatment + Year+ (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Halimeda",])
summary(halimeda_blmer_add)

anova(halimeda_blmer_int, halimeda_blmer_add)
#very sig interaction

#test for sig of fixed effects
halimeda_blmer_add_treat<-blmer(lpercent~Year  +  (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Halimeda",])

halimeda_blmer_add_year<-blmer(lpercent~Treatment  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Halimeda",])

#year
anova(halimeda_blmer_add, halimeda_blmer_add_year)
#very sig decline by year

#treat
anova(halimeda_blmer_add, halimeda_blmer_add_treat)
#ns treatment effect

##test for sig of random effects
halimeda_blmer_null<-blmer(lpercent~1  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Halimeda",])
anova(halimeda_blmer_int, halimeda_blmer_null)
##fixed effects very important

#extract effect size and 95% CI
halimedasumm_int<-summ(halimeda_blmer_int, conf.method="profile",  confint=TRUE,scale=TRUE,  pvals=FALSE)
halimedasumm_int

b<-plot_coefs(halimedasumm_int, inner_ci_level=0.75, ci_levels=0.95, exp=TRUE, color.class=c("black", "black", "black"), conf.method="profile", coefs=c("Treatment:Year","Treatment","Year"))+ 
  ggtitle("Halimeda")+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank()) #remove gridlines
b


r.squaredGLMM(object=halimeda_blmer_int)
#           R2m       R2c
# 0.06684222 0.9232066



#CCA
CCA_blmer_int<-blmer(lpercent~Treatment*Year  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="CCA",])
summary(CCA_blmer_int)

CCA_blmer_add<-blmer(lpercent~Treatment + Year+ (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="CCA",])
summary(CCA_blmer_add)

anova(CCA_blmer_int, CCA_blmer_add)
#very sig interaction

#test for sig of fixed effects
CCA_blmer_add_treat<-blmer(lpercent~Year  +  (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="CCA",])

CCA_blmer_add_year<-blmer(lpercent~Treatment  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="CCA",])

#year
anova(CCA_blmer_add, CCA_blmer_add_year)
#very sig decline by year

#treat
anova(CCA_blmer_add, CCA_blmer_add_treat)
#ns treatment effect

##test for sig of random effects
CCA_blmer_null<-blmer(lpercent~1  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="CCA",])
anova(CCA_blmer_int, CCA_blmer_null)
##fixed effects very important

#extract effect size and 95% CI
CCAsumm_int<-summ(CCA_blmer_int, conf.method="profile",  confint=TRUE,scale=TRUE,  pvals=FALSE)
CCAsumm_int

c<-plot_coefs(CCAsumm_int, inner_ci_level=0.75, ci_levels=0.95, exp=TRUE, color.class=c("black", "black", "black"), conf.method="profile", coefs=c("Treatment:Year","Treatment","Year"))+ 
  ggtitle("CCA")+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank()) #remove gridlines
c


r.squaredGLMM(object=CCA_blmer_int)
#           R2m       R2c
# 0.08199697 0.9315905



#Pavement
Pavement_blmer_int<-blmer(Percent~Treatment*Year  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Pavement",])
summary(Pavement_blmer_int)

Pavement_blmer_add<-blmer(Percent~Treatment + Year+ (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Pavement",])
summary(Pavement_blmer_add)

anova(Pavement_blmer_int, Pavement_blmer_add)
#very sig interaction

#test for sig of fixed effects
Pavement_blmer_add_treat<-blmer(Percent~Year  +  (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Pavement",])

Pavement_blmer_add_year<-blmer(Percent~Treatment  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Pavement",])

#year
anova(Pavement_blmer_add, Pavement_blmer_add_year)

#treat
anova(Pavement_blmer_add, Pavement_blmer_add_treat)


##test for sig of random effects
Pavement_blmer_null<-blmer(Percent~1  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Pavement",])
anova(Pavement_blmer_int, Pavement_blmer_null)
##fixed effects very important


#extract effect size and 95% CI
Pavementsumm_int<-summ(Pavement_blmer_int, conf.method="profile",  confint=TRUE,scale=TRUE,  pvals=FALSE)
Pavementsumm_int

d<-plot_coefs(Pavementsumm_int, inner_ci_level=0.75, ci_levels=0.95, color.class=c("black", "black", "black"), conf.method="profile", coefs=c("Treatment:Year","Treatment","Year"))+ 
  ggtitle("Pavement")+
  xlim(c(-36, 36))+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank()) #remove gridlines
d


r.squaredGLMM(object=Pavement_blmer_int)
#           R2m       R2c
# 0.2786702 0.6973881



#Rubble
Rubble_blmer_int<-blmer(lpercent~Treatment*Year  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Rubble",])
summary(Rubble_blmer_int)

Rubble_blmer_add<-blmer(lpercent~Treatment + Year+ (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Rubble",])
summary(Rubble_blmer_add)

anova(Rubble_blmer_int, Rubble_blmer_add)
#no sig interaction

#test for sig of fixed effects
Rubble_blmer_add_treat<-blmer(lpercent~Year  +  (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Rubble",])

Rubble_blmer_add_year<-blmer(lpercent~Treatment  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Rubble",])

#year
anova(Rubble_blmer_add, Rubble_blmer_add_year)
#ns

#treat
anova(Rubble_blmer_add, Rubble_blmer_add_treat)
#ns treatment effect

##test for sig of random effects
Rubble_blmer_null<-blmer(lpercent~1  + (1|Atoll/Island),  data=benthic_trans_group_meta[benthic_trans_group_meta$Group=="Rubble",])
anova(Rubble_blmer_int, Rubble_blmer_null)
##fixed effects not sig

#extract effect size and 95% CI
Rubblesumm_add<-summ(Rubble_blmer_add, conf.method="profile",  confint=TRUE,scale=TRUE,  pvals=FALSE)
Rubblesumm_add

Rubblesumm_int<-summ(Rubble_blmer_int, conf.method="profile",  confint=TRUE,scale=TRUE,  pvals=FALSE)
Rubblesumm_int

e<-plot_coefs(Rubblesumm_int, inner_ci_level=0.75, ci_levels=0.95, exp=TRUE, color.class=c("black", "black", "black"), conf.method="profile", coefs=c("Treatment:Year","Treatment","Year"))+ 
  ggtitle("Rubble")+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank()) #remove gridlines
e


r.squaredGLMM(object=Rubble_blmer_int)
#           R2m       R2c
# 0.01292026 0.8420805


```


##LMM for structural complexity
```{r}

#check meta-data
uvc_meta
#make year a factor
uvc_meta$Year<-as.factor(uvc_meta$Year)

#look at summarized data
structure_treat <-ddply(uvc_meta, c("Year","Treatment"), summarise,
          Mean_Structure = mean(Structure), 
          sd = sd(Structure),
          n = length(Structure),
          se = sd(Structure)/sqrt(n))

structure_treat
#looks similar across groups

#relevel factor to match other models
uvc_meta$Treatment<-relevel(uvc_meta$Treatment, "Rats")

##Run a model
sc_blmer_int<-blmer(Structure~Treatment*Year  + (1|Atoll/Island),  data=uvc_meta)
summary(sc_blmer_int)

sc_blmer_add<-blmer(Structure~Treatment + Year+ (1|Atoll/Island),  data=uvc_meta)
summary(sc_blmer_add)

anova(sc_blmer_int, sc_blmer_add)
#no sig interaction

#test for sig of fixed effects
sc_blmer_add_treat<-blmer(Structure~Year  +  (1|Atoll/Island),  data=uvc_meta)

sc_blmer_add_year<-blmer(Structure~Treatment  + (1|Atoll/Island),  data=uvc_meta)

#year
anova(sc_blmer_add, sc_blmer_add_year)
#ns

#treat
anova(sc_blmer_add, sc_blmer_add_treat)
#ns treatment effect

##test for sig of random effects
sc_blmer_null<-blmer(Structure~1  + (1|Atoll/Island),  data=uvc_meta)
anova(sc_blmer_int, sc_blmer_null)
##fixed effects not important

#extract effect size and 95% CI
scsumm_add<-summ(sc_blmer_add, conf.method="profile",  confint=TRUE,scale=TRUE,  pvals=FALSE)
scsumm_add

scsumm_int<-summ(sc_blmer_int, conf.method="profile",  confint=TRUE,scale=TRUE,  pvals=FALSE)
scsumm_int

f<-plot_coefs(scsumm_int, inner_ci_level=0.75, ci_levels=0.95,conf.method="profile", coefs=c("Treatment:Year","Treatment","Year"), color.class=c("black", "black", "black"),  point.size = 3)+ 
  ggtitle("Structural complexity")+
  xlim(c(-1.2,1.2))+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank()) #remove gridlines 
f

r.squaredGLMM(object=sc_blmer_int)
#           R2m       R2c
#0.01302966 0.8172735

```

##combine plots
```{r}
library(gridExtra)

b1<-arrangeGrob(a,c,b, d,e,f, nrow=3)

#ggsave("benthos_effect_sizes.eps", b1, units="mm", width=180, height=180, dpi=1200)

```



##Plots of individual benthic groups
```{r, echo = FALSE}
library(ggplot2)
library('gridExtra') 

#get means and se's
benthic_group_treat <-ddply(benthic_trans_group_meta, c("Year","Treatment", "Group"), summarise,
          Mean_Percent = mean(Percent), 
          se = sd(Percent)/sqrt(length(Percent)))
benthic_group_treat
benthic_group_treat$Year<-as.factor(benthic_group_treat$Year)


###plot each FG seperately, then mix together
a<-ggplot(benthic_group_treat[benthic_group_treat$Group=="Hard_Coral",], aes(x=Year, y=Mean_Percent, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Percent-se), ymax=(Mean_Percent+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
   scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("(a) Hard coral") +
  theme_bw() + 
   theme( #axis.text.y = element_blank(), #remove y-axis labels
panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position = "none") + 
        scale_y_continuous(name = "Percent cover", limits = c(0,35), breaks=seq(0,30,10)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching"))  

b<-ggplot(benthic_group_treat[benthic_group_treat$Group=="CCA",], aes(x=Year, y=Mean_Percent, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Percent-se), ymax=(Mean_Percent+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("(b) CCA") +
  theme_bw() + 
    theme( #axis.text.y = element_blank(), #remove y-axis labels
panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        #legend.title = element_blank(), #remove legend title
        #legend.box.background = element_rect(colour = "black", fill='white', linetype='solid', size=1), 
       # legend.justification=c(1,.9), legend.position=c(1,.9)) + #  fix legend
        scale_y_continuous(name = "Percent cover", limits = c(0,35), breaks=seq(0,30,10)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching"))  


c<-ggplot(benthic_group_treat[benthic_group_treat$Group=="Halimeda",], aes(x=Year, y=Mean_Percent, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Percent-se), ymax=(Mean_Percent+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("(c) Halimeda") +
  theme_bw() + 
    theme(#axis.text.y = element_blank(), #remove y-axis labels
  panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Percent cover", limits = c(0,35), breaks=seq(0,30,10)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching"))  
        


d<-ggplot(benthic_group_treat[benthic_group_treat$Group=="Pavement",], aes(x=Year, y=Mean_Percent, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Percent-se), ymax=(Mean_Percent+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                           labels=c("Seabirds", "Rats") ) +
  ggtitle("(d) Pavement") +
  theme_bw() + 
  theme(#axis.text.y = element_blank(), #remove y-axis labels
  panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Percent cover", limits = c(0,56), breaks=seq(0,55,10)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching")) 

e<-ggplot(benthic_group_treat[benthic_group_treat$Group=="Rubble",], aes(x=Year, y=Mean_Percent, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Percent-se), ymax=(Mean_Percent+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                           labels=c("Seabirds", "Rats") ) +
  ggtitle("(e) Rubble") +
  theme_bw() + 
  theme(#axis.text.y = element_blank(), #remove y-axis labels
          panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Percent cover", limits = c(0,15), breaks=seq(0,15,5)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching")) 

#now plot structural complexity
structure_treat

f<-ggplot(structure_treat, aes(x=Year, y=Mean_Structure, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Structure-se), ymax=(Mean_Structure+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                           labels=c("Seabirds", "Rats") ) +
  ggtitle("(f) Structural complexity") +
  theme_bw() + 
  theme(#axis.text.y = element_blank(), #remove y-axis labels
          panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Rugosity index", limits = c(1,6), breaks=seq(1,6,1)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching")) 
f

#show plots
grid.arrange(a, b,c,d,e,f, nrow = 3) 

#export
g <- arrangeGrob(a, b,c,d,e,f, nrow = 3) #generates g
#ggsave("filename.eps", g, units="mm", width=180, height=270, dpi=1200)

```




#Look at hard coral by genus
```{r}

#make new data_frame with only hard coral
head(benthic)
coral_only<-benthic[benthic$Group=="Hard_Coral",]
head(coral_only)
#drop levels
coral_only$Species<-factor(coral_only$Species)
levels(coral_only$Species)

#sum each genus by transect
coral_sp_trans <-ddply(coral_only, c("Year","Island", "Transect", "Species"), summarise,
          Sum_Percent = sum(Percent), 
          .drop=FALSE)
coral_sp_trans

#merge with meta-data
coral_sp_trans_meta<-merge(uvc_meta, coral_sp_trans, c("Year", "Island", "Transect"))

#look at means by treatment and year
coral_sp_treat <-ddply(coral_sp_trans_meta, c("Year","Treatment", "Species"), summarise,
          Mean_Percent = mean(Sum_Percent), 
          sd = sd(Sum_Percent),
          n = length(Sum_Percent),
          se = sd(Sum_Percent)/sqrt(n),
          .drop=FALSE)
coral_sp_treat

#look at means by year
coral_sp_year <-ddply(coral_sp_trans_meta, c("Year","Species"), summarise,
          Mean_Prop = mean(Sum_Percent), 
          sd = sd(Sum_Percent),
          n = length(Sum_Percent),
          se = sd(Sum_Percent)/sqrt(n),
          .drop=FALSE)
coral_sp_year
#Acropora is dominant, and similar between birdy and ratty islands within years

#run stats on Acropra declines
acrop_int<-lmer(Sum_Percent~Treatment*Year + (1|Atoll/Island), data = coral_sp_trans_meta[coral_sp_trans_meta$Species=="Acropora",])
summary(acrop_int)

acrop_add<-lmer(Sum_Percent~Treatment+Year + (1|Atoll/Island), data = coral_sp_trans_meta[coral_sp_trans_meta$Species=="Acropora",])

anova(acrop_int, acrop_add)
#no sig interaction

acrop_treat<-lmer(Sum_Percent~Year + (1|Atoll/Island), data = coral_sp_trans_meta[coral_sp_trans_meta$Species=="Acropora",])

acrop_year<-lmer(Sum_Percent~Treatment + (1|Atoll/Island), data = coral_sp_trans_meta[coral_sp_trans_meta$Species=="Acropora",])

anova(acrop_add, acrop_treat)
#no sig effect of treatment

anova(acrop_add, acrop_year)
#very sig effect of year

#extract effects
summary(acrop_add)
confint(acrop_add, method = "Wald")
```



##Plot coral cover by atoll
```{r}
library(ggplot2)
library('gridExtra') 

#just summed by transect, then averaged by treatment and year
benthic_group2_atoll <-ddply(benthic_trans_group_meta, c("Year","Atoll", "Group"), summarise,
          Mean_Percent = mean(Percent), 
          sd = sd(Percent),
          n = length(Percent),
          se = sd(Percent)/sqrt(n))
benthic_group2_atoll
benthic_group2_atoll$Year<-as.factor(benthic_group2_atoll$Year)

#re-order
benthic_group2_atoll_reorder<-transform(benthic_group2_atoll,
      Atoll=factor(Atoll,levels=c("Salomon","Peros_Banhos","Great_Chagos_Bank")))
###plot all atolls together
hc<-ggplot(benthic_group2_atoll_reorder[benthic_group2_atoll_reorder$Group=="Hard_Coral",], aes(x=Atoll, y=Mean_Percent, fill=Year, colour = Year, group = Year)) + 
          geom_bar(aes(fill=Year), position=position_dodge(width=.9), stat="identity",  colour = 'black') +
          scale_fill_manual(values=c("#00BFC4","#F8766D"),breaks=c("2015", "2018"),
                           labels=c("Pre-bleaching", "Post-bleaching") ) +
          geom_errorbar(aes(ymin=(Mean_Percent-se), ymax=(Mean_Percent+se)),
               width=.2,                    # Width of the error bars
                colour = 'black',  position=position_dodge(.9)) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.title = element_blank(), #remove legend title
        #legend.box.background = element_rect(colour = "black", fill='white', linetype='solid', size=1), 
        legend.justification=c(1,.9), legend.position=c(.8,.8)) + #  fix legend
        scale_y_continuous(name = "Percent cover", expand =c(0,0),limits = c(0,55), breaks=seq(0,50,10)) +
         scale_x_discrete(name ="", 
                    limits=c("Salomon", "Peros_Banhos", "Great_Chagos_Bank"), labels = c("Salomon", "Peros Banhos", 
                                                                                         "Great Chagos Bank")) 
hc
#ggsave("coral_atoll.eps", hc, units="mm", width=180, height=180, dpi=1200)

```
