
##Analyses and figures for Benkwitt, Wilson, & Graham "Seabird nutrient subsidies alter patterns of algal abundance and fish biomass on coral reefs following a bleaching event", Global Change Biology (In Review)

##Fish

#load data
```{r,  include = FALSE}
uvc_fish <- read.csv("seabird_rat_fish_data_2015_2018.csv", header=TRUE)
head(uvc_fish)
```


#merge uvc data with metadata
```{r,  include = FALSE}
#load transect metadata to merge
uvc_meta <- read.csv("seabirds-bleaching/seabird_rat_transect_metadata.csv", header=TRUE)
head(uvc_meta)

##merge dataframes so have transect metadata with benthic data
fish_fg_trans_meta<-merge(uvc_meta, uvc_fish, c("Year", "Island", "Transect"))
fish_fg_trans_meta

```


#NMDS
```{r}
library(reshape2)
library(vegan)

#Aggregate to island-level
fish_fg_island <-ddply(fish_fg_trans_meta, c("Year","Treatment", "Atoll", "Island", "Function"), summarise,
          Mean_Abundance = mean(Sum_Abundance), 
          Mean_Biomass = mean(Sum_Biomass))
fish_fg_island
fish_fg_island$Year<-as.factor(fish_fg_island$Year)

#add column for combined treatment and year
fish_fg_island$Treat_Year<-paste(fish_fg_island$Treatment, fish_fg_island$Year, sep = '_')
#change character to factor
fish_fg_island$Treat_Year<-as.factor(fish_fg_island$Treat_Year)


#convert to wide
fish_fg_island_wide <- dcast(fish_fg_island, Year + Atoll + Island + Treatment + Treat_Year ~ Function, value.var="Mean_Abundance")
fish_fg_island_wide

#log data
log_fish_fg<-log(fish_fg_island_wide[6:ncol(fish_fg_island_wide)]+1)


###RUN NMDS
nmds_fish_fg<-metaMDS(log_fish_fg, distance = "bray", trymax=200, k=2)
nmds_fish_fg
#stress = 0.07028299 - ok
plot(nmds_fish_fg)
scores(nmds_fish_fg, display="species")

#look at plot and centroids
plot(nmds_fish_fg)
ord.fit<-envfit(nmds_fish_fg~Treat_Year, data = fish_fg_island_wide)
ord.fit
plot(ord.fit)
#draw polygons
ordihull(nmds_fish_fg, fish_fg_island_wide$Treat_Year, pty=3)
#looks like some differences among groups, now run statistics


#PERMANOVA
#anosim more influenced by dispersion, so use adonis

#set blocks by atoll
perm_fish <- how(nperm = 999)
setBlocks(perm_fish) <- with(fish_fg_island_wide, Atoll)

#run analysis
ad_fish<-adonis2(log_fish_fg ~ Treatment*Year, data = fish_fg_island_wide, permutations = perm_fish)
ad_fish
#no significant interaction, but year significant

#drop interaction and run additive model to test main effects
ad_fish1<-adonis2(log_fish_fg ~ Treatment+Year, data = fish_fg_island_wide, permutations = perm_fish)
ad_fish1


#test dispersion using PERMDISP
mod_fish <- betadisper(vegdist(log_fish_fg, method = "bray"), fish_fg_island_wide$Treat_Year)
mod_fish


#test for overall differences
anova(mod_fish)

## Permutation test for pairwise comparisons
permutest(mod_fish, pairwise = TRUE)


#extract mean dispersion values (distance to centroid), along with se's and 95% CI
mod_fish.mean<-tapply(mod_fish$distances, fish_fg_island_wide$Treat_Year, mean)
mod_fish.mean
mod_fish.sd<-tapply(mod_fish$distances, fish_fg_island_wide$Treat_Year, sd)
mod_fish.length<-tapply(mod_fish$distances, fish_fg_island_wide$Treat_Year, length)
mod_fish.se<-mod_fish.sd/sqrt(mod_fish.length)
mod_fish.se
mod_fish.ci_low<-mod_fish.mean-(1.96*mod_fish.se)
mod_fish.ci_high<-mod_fish.mean+(1.96*mod_fish.se)
#combine into a dataframe
mod_fish.out<-as.data.frame(cbind(mod_fish.mean,mod_fish.se, mod_fish.ci_low, mod_fish.ci_high ))
mod_fish.out<-cbind(Treat_Year = rownames(mod_fish.out), mod_fish.out)
mod_fish.out


##SIMPER analysis to test effects of species
sim<-simper(log_fish_fg, fish_fg_island_wide$Year, permutations = 999)
summary(sim)

```


##Extract NMDS results and plot
```{r}
library(ggplot2)
library('gridExtra') 

##Extract site and nmds_fg
fg.sc <- scores(nmds_fish_fg, display = "species", shrink = FALSE) 
fg.sc
site.sc <- scores(nmds_fish_fg, display = "sites", shrink = FALSE) 
site.sc
#merge site.sc with metadata from dataframe
fish_island_nmds<-cbind(site.sc, fish_fg_island_wide[1:5])



#extract species scores
species.scores.fish <- as.data.frame(scores(nmds_fish_fg, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores.fish$species <- rownames(species.scores.fish)  # create a column of species, from the rownames of species.scores
head(species.scores.fish)  #look at the data


#get convex hulls for polygons
hull_nr2015<-fish_island_nmds[fish_island_nmds$Treat_Year == "No_rats_2015", ][chull(fish_island_nmds[fish_island_nmds$Treat_Year == "No_rats_2015", ]$NMDS1, fish_island_nmds[fish_island_nmds$Treat_Year == "No_rats_2015", ]$NMDS2), ] 
hull_nr2018<-fish_island_nmds[fish_island_nmds$Treat_Year == "No_rats_2018", ][chull(fish_island_nmds[fish_island_nmds$Treat_Year == "No_rats_2018", ]$NMDS1, fish_island_nmds[fish_island_nmds$Treat_Year == "No_rats_2018", ]$NMDS2), ] 
hull_r2018<-fish_island_nmds[fish_island_nmds$Treat_Year == "Rats_2018", ][chull(fish_island_nmds[fish_island_nmds$Treat_Year == "Rats_2018", ]$NMDS1, fish_island_nmds[fish_island_nmds$Treat_Year == "Rats_2018", ]$NMDS2), ] 
hull_r2015<-fish_island_nmds[fish_island_nmds$Treat_Year == "Rats_2015", ][chull(fish_island_nmds[fish_island_nmds$Treat_Year == "Rats_2015", ]$NMDS1, fish_island_nmds[fish_island_nmds$Treat_Year == "Rats_2015", ]$NMDS2), ] 
hull.data.fish <- rbind(hull_r2015, hull_nr2015,  hull_r2018,hull_nr2018 )  #combine grp.a and grp.b
hull.data.fish

#get centroids
ord.fit

nmds_f<-
  ggplot() + 
 geom_polygon(data=hull.data.fish[hull.data.fish$Treatment=="Rats",],aes(x=NMDS1,y=NMDS2,fill=Treatment,group=Treat_Year), colour = 'black', alpha = 0.15) + # add the convex hulls - two steps because the blue is fainter
   geom_polygon(data=hull.data.fish[hull.data.fish$Treatment=="No_rats",],aes(x=NMDS1,y=NMDS2,fill=Treatment,group=Treat_Year), colour = 'black', alpha = 0.2) + # add the convex hulls - two steps because the blue is fainter
  geom_text(data=species.scores.fish,aes(x=NMDS1,y=NMDS2,label=species)) +  # add the species labels
  geom_point(data=fish_island_nmds,aes(x=NMDS1,y=NMDS2,shape=Year,colour=Treatment, fill = Treatment), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                           labels=c("Seabirds", "Rats") ) +
    geom_segment(aes(x = 0.1291, y = -0.0112, xend = -0.0529, yend = 0.1019), colour = "#56B4E9", size = 1.2, arrow =   arrow(length = unit(0.2, "inches")))+ #norats
  geom_segment(aes(x = 0.0877, y = 0.0087, xend = -0.2072, yend = -0.0990), colour = "#e41a1c", size = 1.2, arrow =   arrow(length = unit(0.2, "inches")))+ #rats
  scale_x_continuous(limits = c(-0.42, 0.41), breaks = seq(-0.4, 0.4, 0.2)) +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none")


##Make a plot of dispersion means +/- SE
mod_fish.out #from betadisper, above
#create new columns for Treament and year
mod_fish.out$Treatment<-c("No_rats","No_rats","Rats", "Rats")
mod_fish.out$Year<-c("2015","2018","2015", "2018")
mod_fish.out

fishd<-ggplot(mod_fish.out, aes(x=Year, y=mod_fish.mean, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(mod_fish.mean-mod_fish.se), ymax=(mod_fish.mean+mod_fish.se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("(b) Fish") +
  theme_bw() + 
    theme(#axis.text.y = element_blank(), #remove y-axis labels
  panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Dispersion", limits = c(0,0.2), breaks=seq(0,0.2,0.05)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching"))  

fishd

#arrange and save (combined with benthos plots from benthic_analysis.Rmd)
#show plots
grid.arrange(nmds_b, nmds_f, nrow = 2) 
g <- arrangeGrob(nmds_b, nmds_f, nrow = 2) #generates g
#ggsave("filename.eps", g, units="mm", width=180, height=180, dpi=1200)

```





#LMMs for individual groups
```{r}
library(lme4) #lmms
library(blme)
library(ggplot2)
library(jtools) #plot_summ
library(MuMIn) #r.squaredGLMM
  
#look at data
fish_fg_trans_meta
#change year to factor
fish_fg_trans_meta$Year<-as.factor(fish_fg_trans_meta$Year)
#relevel factor
fish_fg_trans_meta$Treatment<-relevel(fish_fg_trans_meta$Treatment, "Rats")

#create column for logged biomass to make coding easier
fish_fg_trans_meta$lbiomass<-log(fish_fg_trans_meta$Sum_Biomass+1)

#run final models, log-transformed when necessary to meet assumptions

####Herbivore 
herb_blmer_int<-blmer(Sum_Biomass~Treatment*Year+Coral_cover+Structure  + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Herbivore",])
summary(herb_blmer_int)

herb_blmer_add<-blmer(Sum_Biomass~Treatment + Year+Coral_cover+Structure+ (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Herbivore",])
summary(herb_blmer_add)

anova(herb_blmer_int, herb_blmer_add)
#no sig interaction

#test for sig of fixed effects
herb_blmer_add_treat<-blmer(Sum_Biomass~Year +Coral_cover+Structure +  (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Herbivore",])

herb_blmer_add_year<-blmer(Sum_Biomass~Treatment +Coral_cover+Structure + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Herbivore",])

#year
anova(herb_blmer_add, herb_blmer_add_year)
#no sig decline by year

#treat
anova(herb_blmer_add, herb_blmer_add_treat)
#sig treatment effect

##test for sig of random effects
herb_blmer_null<-blmer(Sum_Biomass~1  + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Herbivore",])
anova(herb_blmer_int, herb_blmer_null)


#extract effect size and 95% CI
herbsumm_add<-summ(herb_blmer_add, conf.method="profile",  confint=TRUE,scale=TRUE, pvals=FALSE)
herbsumm_add

herbsumm_int<-summ(herb_blmer_int, conf.method="profile",  confint=TRUE,scale=TRUE,  pvals=FALSE)
herbsumm_int

a<-plot_coefs(herbsumm_int, inner_ci_level=0.75, ci_levels=0.95, conf.method="profile",color.class=c("black", "black", "black"), coefs=c("Treatment:Year","Treatment","Year", "Coral_cover", "Structure"))+ 
  ggtitle("Herbivore")+
  xlim(c(-500, 500))+
  theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank()) #remove gridlines
a


r.squaredGLMM(object=herb_blmer_int)
#           R2m       R2c
# 0.1374722 0.5896944



####Piscivore
Pisc_blmer_int<-blmer(Sum_Biomass~Treatment*Year+Coral_cover+Structure  + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Piscivore",])
summary(Pisc_blmer_int)
plot(Pisc_blmer_int)

Pisc_blmer_add<-blmer(Sum_Biomass~Treatment + Year+Coral_cover+Structure+ (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Piscivore",])
summary(Pisc_blmer_add)

anova(Pisc_blmer_int, Pisc_blmer_add)
#no sig interaction

#test for sig of fixed effects
Pisc_blmer_add_treat<-blmer(Sum_Biomass~Year +Coral_cover+Structure +  (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Piscivore",])

Pisc_blmer_add_year<-blmer(Sum_Biomass~Treatment +Coral_cover+Structure + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Piscivore",])

#year
anova(Pisc_blmer_add, Pisc_blmer_add_year)
#no sig decline by year

#treat
anova(Pisc_blmer_add, Pisc_blmer_add_treat)
#sig treatment effect

##test for sig of random effects
Pisc_blmer_null<-blmer(Sum_Biomass~1  + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Piscivore",])
anova(Pisc_blmer_int, Pisc_blmer_null)
##ns 0.12

#extract effect size and 95% CI
Piscsumm_add<-summ(Pisc_blmer_add, conf.method="profile",  confint=TRUE,  scale=TRUE,pvals=FALSE)
Piscsumm_add

Piscsumm_int<-summ(Pisc_blmer_int, conf.method="profile",  confint=TRUE,  scale=TRUE,pvals=FALSE)
Piscsumm_int

b<-plot_coefs(Piscsumm_int, inner_ci_level=0.75, ci_levels=0.95,conf.method="profile",color.class=c("black", "black", "black"), coefs=c("Treatment:Year","Treatment","Year", "Coral_cover", "Structure"))+ 
  ggtitle("Piscivore") +
    xlim(c(-160, 160))+
 theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank()) #remove gridlines

b

r.squaredGLMM(object=Pisc_blmer_int)
#           R2m       R2c
# 0.04336649 0.4488459



####Invertivore
Invert_blmer_int<-blmer(Sum_Biomass~Treatment*Year+Coral_cover+Structure  + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Invertivore",])
summary(Invert_blmer_int)
plot(Invert_blmer_int)

Invert_blmer_add<-blmer(Sum_Biomass~Treatment + Year+Coral_cover+Structure+ (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Invertivore",])
summary(Invert_blmer_add)

anova(Invert_blmer_int, Invert_blmer_add)
#no sig interaction

#test for sig of fixed effects
Invert_blmer_add_treat<-blmer(Sum_Biomass~Year +Coral_cover+Structure +  (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Invertivore",])

Invert_blmer_add_year<-blmer(Sum_Biomass~Treatment +Coral_cover+Structure + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Invertivore",])

#year
anova(Invert_blmer_add, Invert_blmer_add_year)
#no sig decline by year

#treat
anova(Invert_blmer_add, Invert_blmer_add_treat)


##test for sig of random effects
Invert_blmer_null<-blmer(Sum_Biomass~1  + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Invertivore",])
anova(Invert_blmer_int, Invert_blmer_null)


#extract effect size and 95% CI
Invertsumm_add<-summ(Invert_blmer_add, conf.method="profile",  confint=TRUE,  pvals=FALSE)
Invertsumm_add

Invertsumm_int<-summ(Invert_blmer_int, conf.method="profile",  confint=TRUE, scale=TRUE, pvals=FALSE)
Invertsumm_int

c<-plot_coefs(Invertsumm_int, inner_ci_level=0.75, ci_levels=0.95, conf.method="profile", color.class=c("black", "black", "black"), coefs=c("Treatment:Year","Treatment","Year", "Coral_cover", "Structure"))+ 
  ggtitle("Invertivore") + 
  xlim(c(-100, 100))+
 theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank()) #remove gridlines
c


r.squaredGLMM(object=Invert_blmer_int)
#           R2m       R2c
# 0.06964612 0.3071322



####Mixed-diet 
Mix_blmer_int<-blmer(lbiomass~Treatment*Year+Coral_cover+Structure  + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Mixed-diet",])
summary(Mix_blmer_int)
plot(Mix_blmer_int) #outliers, but not terrible without log........

Mix_blmer_add<-blmer(lbiomass~Treatment + Year+Coral_cover+Structure+ (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Mixed-diet",])
summary(Mix_blmer_add)

anova(Mix_blmer_int, Mix_blmer_add)
#no sig interaction

#test for sig of fixed effects
Mix_blmer_add_treat<-blmer(lbiomass~Year +Coral_cover+Structure +  (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Mixed-diet",])

Mix_blmer_add_year<-blmer(lbiomass~Treatment +Coral_cover+Structure + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Mixed-diet",])

#year
anova(Mix_blmer_add, Mix_blmer_add_year)
#no sig decline by year

#treat
anova(Mix_blmer_add, Mix_blmer_add_treat)
#no sig treatment effect


##test for sig of random effects
Mix_blmer_null<-blmer(lbiomass~1  + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Mixed-diet",])
anova(Mix_blmer_int, Mix_blmer_null)
##not sig 

#extract effect size and 95% CI
Mixsumm_add<-summ(Mix_blmer_add, conf.method="profile",  confint=TRUE,  pvals=FALSE)
Mixsumm_add

Mixsumm_int<-summ(Mix_blmer_int, conf.method="profile",  confint=TRUE, scale=TRUE, pvals=FALSE)
Mixsumm_int

d<-plot_coefs(Mixsumm_int, inner_ci_level=0.75, ci_levels=0.95, exp=TRUE, conf.method="profile", color.class=c("black", "black", "black"), coefs=c("Treatment:Year","Treatment","Year", "Coral_cover", "Structure"))+ 
  ggtitle("Mixed-diet")+
# xlim(c(-100, 100))+
 theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank()) #remove gridlines

d

r.squaredGLMM(object=Mix_blmer_int)
#           R2m       R2c
# 0.04246732 0.8084706




####Planktivore 
Plankt_blmer_int<-blmer(lbiomass~Treatment*Year+Coral_cover+Structure  + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Planktivore",])
summary(Plankt_blmer_int)
plot(Plankt_blmer_int)

Plankt_blmer_add<-blmer(lbiomass~Treatment + Year+Coral_cover+Structure+ (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Planktivore",])
summary(Plankt_blmer_add)

anova(Plankt_blmer_int, Plankt_blmer_add)
#no sig interaction

#test for sig of fixed effects
Plankt_blmer_add_treat<-blmer(lbiomass~Year +Coral_cover+Structure +  (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Planktivore",])

Plankt_blmer_add_year<-blmer(lbiomass~Treatment +Coral_cover+Structure + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Planktivore",])

#year
anova(Plankt_blmer_add, Plankt_blmer_add_year)
#sig decline by year

#treat
anova(Plankt_blmer_add, Plankt_blmer_add_treat)
#marginally sig treatment effect


##test for sig of random effects
Plankt_blmer_null<-blmer(lbiomass~1  + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Planktivore",])
anova(Plankt_blmer_int, Plankt_blmer_null)


#extract effect size and 95% CI
Planktsumm_add<-summ(Plankt_blmer_add, conf.method="profile",  confint=TRUE,  pvals=FALSE)
Planktsumm_add

Planktsumm_int<-summ(Plankt_blmer_int, conf.method="profile",  confint=TRUE, scale=TRUE, pvals=FALSE)
Planktsumm_int

e<-plot_coefs(Planktsumm_int, inner_ci_level=0.75, ci_levels=0.95, exp=TRUE, conf.method="profile", color.class=c("black", "black", "black"), coefs=c("Treatment:Year","Treatment","Year", "Coral_cover", "Structure"))+ 
  ggtitle("Planktivore") +
theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank()) #remove gridlines

e


r.squaredGLMM(object=Plankt_blmer_int)
#           R2m       R2c
# 0.03632118 0.7828376




####Corallivore 
Corall_blmer_int<-blmer(lbiomass~Treatment*Year+Coral_cover+Structure  + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Corallivore",])
summary(Corall_blmer_int)
plot(Corall_blmer_int)

Corall_blmer_add<-blmer(lbiomass~Treatment + Year+Coral_cover+Structure+ (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Corallivore",])
summary(Corall_blmer_add)

anova(Corall_blmer_int, Corall_blmer_add)
#no sig interaction

#test for sig of fixed effects
Corall_blmer_add_treat<-blmer(lbiomass~Year +Coral_cover+Structure +  (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Corallivore",])

Corall_blmer_add_year<-blmer(lbiomass~Treatment +Coral_cover+Structure + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Corallivore",])

#year
anova(Corall_blmer_add, Corall_blmer_add_year)
#SIGNIFICANT

#treat
anova(Corall_blmer_add, Corall_blmer_add_treat)
#ns


##test for sig of random effects
Corall_blmer_null<-blmer(lbiomass~1  + (1|Atoll/Island),  data=fish_fg_trans_meta[fish_fg_trans_meta$Function=="Corallivore",])
anova(Corall_blmer_int, Corall_blmer_null)
##very sig

#extract effect size and 95% CI
Corallsumm_add<-summ(Corall_blmer_add, conf.method="profile",  confint=TRUE,  pvals=FALSE)
Corallsumm_add

Corallsumm_int<-summ(Corall_blmer_int, conf.method="profile",  confint=TRUE, scale=TRUE, pvals=FALSE)
Corallsumm_int

f<-plot_coefs(Corallsumm_int, inner_ci_level=0.75, ci_levels=0.95,exp=TRUE, conf.method="profile", color.class=c("black", "black", "black"), coefs=c("Treatment:Year","Treatment","Year", "Coral_cover", "Structure"))+ 
  ggtitle("Corallivore") +
theme(panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank()) #remove gridlines

f



r.squaredGLMM(object=Corall_blmer_int)
#           R2m       R2c
# 0.05262509 0.8293092
```


##bring effect size plots together
```{r}

library(gridExtra)

f1<-grid.arrange(a,b,c,d,e,f, nrow=3)

#ggsave("fish_effect_sizes.eps", f1, units="mm", width=180, height=180, dpi=1200)


```


##Plots of individual fish groups
```{r, echo = FALSE}
library(ggplot2)
library('gridExtra') 

#get means and se's
fish_uvc_treat_sum <-ddply(fish_fg_trans_meta, c("Year","Treatment", "Function"), summarise,
          Mean_Biomass = mean(Sum_Biomass), 
          se = sd(Sum_Biomass)/sqrt(length(Sum_Biomass)))
fish_uvc_treat_sum

fish_uvc_treat_sum$Year<-as.factor(fish_uvc_treat_sum$Year)

###plot each FG seperately, then mix together.
a<-ggplot(fish_uvc_treat_sum[fish_uvc_treat_sum$Function=="Herbivore",], aes(x=Year, y=Mean_Biomass, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Biomass-se), ymax=(Mean_Biomass+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
    scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("Herbivore") +
  theme_bw() + 
  theme(#axis.text.y = element_blank(), #remove y-axis labels
          panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
                legend.position="none") +
       scale_y_continuous(name = "Biomass (kg/ha)", limits = c(0,650), breaks=seq(0,600,200)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching"))


b<-ggplot(fish_uvc_treat_sum[fish_uvc_treat_sum$Function=="Piscivore",], aes(x=Year, y=Mean_Biomass, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Biomass-se), ymax=(Mean_Biomass+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
        scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("Piscivore") +
  theme_bw() + 
  theme(#axis.text.y = element_blank(), #remove y-axis labels
          panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
       scale_y_continuous(name = "Biomass (kg/ha)",limits = c(0,165), breaks=seq(0,150,50)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching")) 


c<-ggplot(fish_uvc_treat_sum[fish_uvc_treat_sum$Function=="Invertivore",], aes(x=Year, y=Mean_Biomass, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Biomass-se), ymax=(Mean_Biomass+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
       scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("Invertivore") +
  theme_bw() + 
  theme(#axis.text.y = element_blank(), #remove y-axis labels
          panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
       scale_y_continuous(name = "Biomass (kg/ha)", limits = c(0,100), breaks=seq(0,100,20)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching")) 


d<-ggplot(fish_uvc_treat_sum[fish_uvc_treat_sum$Function=="Mixed-diet",], aes(x=Year, y=Mean_Biomass, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Biomass-se), ymax=(Mean_Biomass+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
       scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("Mixed-diet") +
  theme_bw() + 
  theme(#axis.text.y = element_blank(), #remove y-axis labels
          panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
         scale_y_continuous(name = "Biomass (kg/ha)", limits = c(0,50), breaks=seq(0,50,10)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching")) 


e<-ggplot(fish_uvc_treat_sum[fish_uvc_treat_sum$Function=="Planktivore",], aes(x=Year, y=Mean_Biomass, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Biomass-se), ymax=(Mean_Biomass+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
       scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("Planktivore") +
  theme_bw() + 
  theme(#axis.text.y = element_blank(), #remove y-axis labels
          panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Biomass (kg/ha)", limits = c(0,40), breaks=seq(0,40,10)) +
      scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching"))  

f<-ggplot(fish_uvc_treat_sum[fish_uvc_treat_sum$Function=="Corallivore",], aes(x=Year, y=Mean_Biomass, fill=Treatment, colour = Treatment, shape = Year, group = Treatment)) + 
            geom_line(colour = 'black') +
            geom_errorbar(aes(ymin=(Mean_Biomass-se), ymax=(Mean_Biomass+se)),
               width=.2,                    # Width of the error bars
                colour = 'black') +
          geom_point(aes(fill=Treatment, shape = Year), position=position_dodge(width=0), stat="identity", size=5,  colour = 'black') +
       scale_shape_manual(values=c(21,24),breaks=c("2015", "2018"),
                            labels=c("Pre-bleaching", "Post-bleaching") ) +
   scale_fill_manual(values=c("#56B4E9","#e41a1c"),breaks=c("No_rats", "Rats"),
                            labels=c("Seabirds", "Rats") ) +
  ggtitle("Corallivore") +
  theme_bw() + 
  theme(#axis.text.y = element_blank(), #remove y-axis labels
          panel.grid.major = element_blank(), # remove gridlines
        panel.grid.minor = element_blank(), #remove gridlines
        strip.background = element_blank(), #remove gray boxes
        text = element_text(size=20),
        legend.position="none") +
        scale_y_continuous(name = "Biomass (kg/ha)", limits = c(0.00,10), breaks=seq(0,10,2)) +
        scale_x_discrete(name ="", 
                    limits=c("2015", "2018"), labels = c("Pre-bleaching", "Post-bleaching"))

#export plot
grid.arrange(a, b, c, d, e, f, nrow = 3) 
fish_plot <- arrangeGrob(a, b,c,d,e,f, nrow = 3) #generates plot for export
#ggsave("filename.eps", fish_plot, units="mm", width=180, height=270, dpi=1200)

```

